A Comparison of Parallel Graph Processing Benchmarks

### 2017-02-17 papers/publication/paper-pollard.tex
conclusion

Unlike these approaches, we focus on the \emph{depth} of analysis rather than the breadth via diversity of datasets or graph processing systems. We analyze  power and energy consumption in addition to details such as the number of iterations or the time to construct the data structures. Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for timing).

@@

We select BFS because the canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}. The advantage of the Graph500 is it provides standardized measurement specifications and dataset generation. The primary drawback with the Graph500 is it measures a single algorithm.

@@

PageRank is selected because of its popularity; most libraries provide reference implementations. One challenge with using PageRank is the stopping criterion; this is addressed in Sect.~\ref{sec:perf}. Verification of the PageRanks is beyond the scope of this paper though this may explain some of the large performance discrepancies.

This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be easily extended to other algorithms.

\begin{table}
	\centering
	\caption{The operating system is GNU/Linux version 4.4.0-22. The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit. We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
	% For arya I deleted Max RAM Freq	2133MHz
	\pgfplotstabletypeset[
	header=false,
	col sep=tab,
	string type,
	every head row/.style={output empty row, before row=\bottomrule},
	columns/0/.style={column type={|r|}},
	columns/1/.style={column type={l|}},
	every last row/.style={after row=\toprule},
	]{../report/specs.csv}
	\label{tab:specs}
\end{table}

@@

We attempt to use similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon. This could account for the increased number of iterations. We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criteria, where $i$ is the iteration and $n$ is the number of vertices. We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating point number. This is done with the goal of making the stoppping criteria as similar as possible.

Furthermore, the logarithmic scale in Figs.~\ref{fig:pr} and~\ref{fig:bfs-time} compresses the apparent variance in runtime. Each platform in Fig.~\ref{fig:pr} has a relative standard deviation between $1/4$ and $1/2$ that of the same system executing SSSP.

@@

\section{Future Work and Conclusion}
We have presented an approach to analyze parallel graph processing performance at fine detail, both measuring power consumption and the two fundamental phases of execution. Still, the problem of choosing a system for graph processing at large scale is far from being simple. We make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation. Our results required minor changes to the original projects to add calls to PAPI sampling and those changes are also available on forked versions of the repositories.

Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets and the most scalable. However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices. It should be noted that GAP is also the most recent of projects. Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite to as a good test of performance.

There are a vast number of directions this work could go. To ensure fairness, each platform must be configured to use the same graphs and the same roots. In the case of GraphBIG this required the file to be loaded in for each experiment. This was by far the most time consuming of the experiments; the file reading for GraphBIG being done serially on an uncompressed ASCII text format for graphs. This limited the sizes of the experiments we could run. Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail \cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question.

>>>>

Unlike these approaches, we focus on the \emph{depth} of analysis rather than the breadth via diversity of datasets or graph processing systems. We analyze  power and energy consumption in addition to details such as the number of iterations or the time to construct the data structures. Moreover, our approach requires little knowledge of the inner workings of each system; the data are collected using either hardware counter sampling of model-specific registers (for power) or through parsing log files (for execution time).

@@

We select BFS because the canonical performance leaderboard for parallel graph processing is the Graph500 \cite{Murphy:2010:Graph500}. The advantage of the Graph500 is that it provides standardized measurement specifications and dataset generation. The primary drawback with the Graph500 is it measures a single algorithm.

@@

PageRank is selected because of its popularity; most libraries provide reference implementations. One challenge with using PageRank is the stopping criterion; this is addressed in Sect.~\ref{sec:perf}. Verification of the PageRank results is beyond the scope of this paper, although this may explain some of the large performance discrepancies.

This approach is not specific to a particular algorithm; measuring the execution time, data structure construction time, and power consumption can be applied easily to other algorithms.

\begin{table}
	\centering
	\caption{The operating system is GNU/Linux version 4.4.0-22. The disparity between the CPU's advertised clock speed and the ``CPU Clock'' row is a result of the Turbo Boost technology which can increase the clock speed to a limit. We use the manufacturer's published maximum clock speeds which can be found at \url{http://ark.intel.com}.}
	% For arya I deleted Max RAM Freq	2133MHz
	\pgfplotstabletypeset[
	header=false,
	col sep=tab,
	string type,
	every head row/.style={output empty row, before row=\bottomrule},
	columns/0/.style={column type={|p{1in}|}},
	columns/1/.style={column type={p{3.2in}|}},
	every last row/.style={after row=\toprule},
	]{../report/specs.csv}
	\label{tab:specs}
\end{table}

@@

We attempt to use similar stopping criteria for each system, but GraphMat executes until no vertices change rank; effectively its stopping criterion requires the $\infty$-norm be less than machine epsilon. This could account for the increased number of iterations. We adjusted the other systems to use $\sum_{k=1}^{n} |p_k^{(i)} - p_k^{(i-1)}| < \epsilon $ as the stopping criteria, where $i$ is the iteration and $n$ is the number of vertices. We use $\epsilon = 6 \times 10^{-8}$ because this value is approximately machine epsilon for a single precision floating point number. This is done with the goal of making the stopping criteria as similar as possible.

Moreover, the logarithmic scale in Figures~\ref{fig:bfs-time} and~\ref{fig:pr} compresses the apparent variance in runtime. Each platform in Fig.~\ref{fig:pr} has a relative standard deviation between $1/4$ and $1/2$ that of the same system executing SSSP.

@@

\section{Future Work and Conclusion}
We have presented a systematic approach to analyzing parallel graph processing performance, both measuring power consumption and the two fundamental phases of execution. While the comparison presented here can help choose among alternatives for the selected packages and algorithms, the problem of choosing a system for graph processing at large scale is far from simple. Hence, we make our code available at \url{https://github.com/HPCL/easy-parallel-graph} and encourage further experimentation. Our results required minor changes to the original projects to add calls to PAPI sampling and those changes are also available on forked versions of the repositories.

Overall, the GAP Benchmark Suite was the highest-performing system across the given datasets and the most scalable. However, this is only for relatively small graphs; all the graphs used here had at most $2^{22}$ vertices. It should be noted that GAP is also the most recent of projects. Thus, we recommend graph processing algorithm designers compare their implementations against the GAP Benchmark Suite as a good test of performance.

Our work can continue in many directions. To ensure fairness, each platform must be configured to use the same graphs and the same roots. In the case of GraphBIG, this required the file to be loaded in for each experiment. By far the most time consuming of the experiments was GraphBIG's serial file input of uncompressed ASCII text format for graphs. This limited the sizes of the experiments we could run. Graphalytics also had circumstances with the more computationally expensive algorithms  where certain experiments fail \cite{Iosup:2016:Graphalyticstech}, so determining whether an algorithm will finish given a particular machine, input size, runtime limit, and resources is an important unanswered question we plan to pursue further.
